Parsing Postfix Log Files

### 2009-08-27 doc/logparser-implementation.tex
Changes made to improve appearance after switching to one-and-a-half spacing.
Changes to capitalisation of titles, as specified by Howard Hamilton. Fixed that/than type identified by Joe Carthy.
Turned regex into an acronym everywhere.

The next three sections document the implementation of the three components
of the architecture.  First is the framework, including its initialisation
phase, the parsing process, and the conveniences it offers to users of
\parsername{}; that section finishes by describing features that are
important for this thesis but not strictly necessary for parsing: the
performance data collected by the framework, the optimisations that can be
disabled to show their effect, and the debugging options the framework
provides.  The second component of the architecture is the actions,
starting with a graph showing how often each action is specified by rules,
a description of why some actions are more popular than others, and how
this popularity has no influence on how often an action is invoked.  Every
action that is part of \parsername{} is documented, and the actions section
concludes with a description of the process of adding a new action.  Rules
are the third component of the architecture, and also the most visible to
advanced users, e.g.\ systems administrators, because it is likely that
they will need to add their own rules to recognise their own log lines.  A
sample rule used by \parsername{} is examined, with every field clearly
documented; that is followed by a description of adding new rules, and how
to determine the value of each of the new rule's fields.  \parsername{}
provides a utility to create regexes from unrecognised log lines; the
algorithm it uses is documented, including the differences between it and
the original algorithm it is based on.  The rules section finishes with a
discussion of how \parsername{} uses rule conditions and overlapping rules,
plus a description of the regex snippets provided to ease the process of
writing regexes.

@@

The database schema can be conceptually divided in two: the rules used to
recognise log lines, and the data saved from the parsing of log files.
Each rule has a regex to recognise log lines and capture data from them,
and specifies the action to be invoked when a log line is recognised; they
also have several other fields used by the parser, and several fields that
aid the user in understanding the meaning of the log lines recognised by
each rule.  The rules are described in detail in \sectionref{rules in
implementation}, but the rules table is documented in \sectionref{rules
table} with the rest of the database schema.

@@

\subsection{Using A Database To Provide An Application Programming Interface}

@@

    \item [regex] The regex to recognise log lines with, as documented in
        \sectionref{regex components}.

@@

    \item That the regex is valid.  The regex will have regex components
        expanded (see \sectionref{regex components}), and will also be
        compiled for efficiency (\sectionref{Caching compiled regexes}).

    \item For overlap between the data captured by the regex and additional
        data specified in either \texttt{connection\_data} or
        \texttt{result\_data}.

    \item That \texttt{connection\_data}, \texttt{result\_data}, and the
        regex captures specify valid fields to save data to.

@@

    \item Each regex is compiled and the result cached when the ruleset is
        loaded; this optimisation can be disabled and the regex compiled
        every time it is used when recognising log lines.  The effect this
        optimisation has is addressed shown in \sectionref{Caching compiled
        regexes}.

@@

    \item Every time a log line is recognised, the recognising rule's regex
        and the log line can be printed, so that the user can verify that
        the correct rule recognised each log line.  This is equivalent to
        setting every rule's debug field to true.

@@

\subsection{Description Of Each Action}

@@

If the log line has enough information to identify the correct connection
or mail, each action will save all the data captured by the recognising
rule's regex to the connection or mail; usually, log lines that lack
identifying information will be processed by the \action{UNINTERESTING}
action.  Each action is passed the same arguments:

@@

    \item [data] The data captured from the log line by the rule's regex.

@@

    \item [CLEANUP\_PROCESSING] \daemon{cleanup} processes every mail that
        passes through Postfix; details of what it does are available in
        \sectionref{Postfix Daemons}.  This action saves all data captured
        by the rule's regex if the log line has not come after a timeout
        (see \sectionref{discarding cleanup log lines}); it also creates
        the mail if necessary, setting its unknown origin flag (see
        \sectionref{pickup logging after cleanup}).

@@

    \item [DELIVERY\_REJECTED] Postfix rejected a mail delivery attempt
        from the remote client.  This is the action most frequently
        specified by rules, because so many different restrictions are used
        to reject delivery attempts.  This action is quite simple: if the
        log line contains a queueid, save the data captured by the rule's
        regex to the mail identified by that queueid; otherwise save it to
        the connection identified by the \acronym{pid} in the log line.

@@

    \item [MAIL\_BOUNCED] This action behaves in exactly the same way as
        the \action{SAVE\_DATA} action; it saves all data captured by the
        recognising rule's regex, and does nothing more.  It is a separate
        action to distinguish delivery attempts that bounce from other
        delivery attempts.

@@

    \item [MAIL\_SENT] This action behaves in exactly the same way as the
        \action{SAVE\_DATA} action; it saves all data captured by the
        recognising rule's regex, and does nothing more.  It is a separate
        action to distinguish successful delivery attempts from other
        delivery attempts.

@@

    \item [SAVE\_DATA] Every action that can locate the correct data
        structure in the state tables saves any data captured by the
        recognising rule's regex to it.  The \action{SAVE\_DATA} saves data
        in this way but does not do anything else; it is invoked for log
        lines that contain useful data but do not require any further
        processing.

@@

The role of the rules in the architecture is covered in detail in
\sectionref{rules in architecture}; this section is concerned with the
practical aspects of how rules are implemented and used in \parsername{}.
The structure of the rules table has already been documented in
\sectionref{rules table}; that description will not be repeated here, but
should be fleshed out by the sample rule in \sectionref{example rule in
implementation}.  The process of creating new rules from unparsed log lines
is dealt with in \sectionref{creating new rules in implementation},
followed by the algorithm used by the utility supplied with \parsername{}
that creates regexes from unparsed log lines; the regex components provided
by \parsername{} to ease writing of complex regexes are covered in
\sectionref{regex components}.  The rule conditions used in \parsername{}
are the penultimate topic (\sectionref{rule conditions in implementation}),
and this section concludes with some suggestions for how to detect
overlapping rules.

@@

    \item [program and regex:] program is used to restrict the log lines
        this rule will be used to recognise; see \sectionref{rule
        conditions in implementation} for details.  The regex does the
        actual recognition of log lines, and data captured by the regex
        (e.g.\ sender, recipient) will be automatically saved to the
        results and connections tables.  The regex components used in the
        regex are described in \sectionref{regex components}.

@@

To add a new rule a new row must be added to the rules table in the
database, containing all the required attributes: action, name,
description, program, and regex; all the other attributes are either
optional, are set by the parser, or have sensible defaults.  The name and
description fields should be set based on the meaning of the log line, to
help others understand which log lines this rule will recognise; the value
for the program field will be obvious from the unrecognised log lines.  The
action depends on the what the log line represents, e.g.\ a delivery
rejection, a mail being delivered, some useful information, or something
else; examine the list of actions in \sectionref{actions in detail in
implementation} to determine the correct one.  The regex needs to be based
on the log line, but see below for a tool to ease the process of creating
regexes from unrecognised log lines.

@@

A program is provided with \parsername{} to ease the process of creating
new rules from unrecognised log lines, based on the algorithm developed by
Risto Vaarandi for his \acronym{SLCT}~\cite{slct-paper}.  The differences
between the two algorithms will be outlined as part of the explanation
below.  The core of the \acronym{SLCT} algorithm is quite simple: programs
generally create log lines by substituting variable words into a fixed
pattern, and analysis of the frequency with which each word occurs can be
used to determine whether the word is variable or part of the fixed
pattern.  This classification can be used to group similar log lines and
generate a regex to match each group of log lines.  The algorithm has five
steps:

@@

    \item [Pre-process the input.]  The modified algorithm begins by
        leveraging the knowledge gained from writing rules and developing
        \parsername{}; it performs many substitutions on the input log
        lines, replacing commonly occurring variable terms (e.g.\ email
        addresses, IP addresses, the standard start of rejection messages)
        with keywords described in \sectionref{regex components}.  The
        purpose of this step is to utilise existing knowledge to create
        more accurate regexes; it replaces many variable words with fixed
        words, improving the subsequent classification of words as fixed or
        variable.  Regex metacharacters in the log line will be escaped, to
        avoid generating invalid or incorrect regexes.  The altered log
        lines are written to a temporary file, which the next stage will
        use instead of the original input file.

@@

    \item [Classify words based on their frequency.]  The frequency of each
        \texttt{(word, word's position within the log line)} tuple is
        checked: if its frequency is greater than the threshold supplied by
        the user (1\% of all log lines is generally a good starting point),
        it is classified as a fixed word, otherwise it is classified as a
        variable term.  If a variable term appears sufficiently often it
        will be misclassified as a fixed term, but that should be noticed
        by the user when reviewing the new regexes, and will be obvious
        when the new rules do not recognise some log lines they are
        expected to.  Variable terms are replaced by \texttt{.+} to match
        one or more of any character.

    \item [Build regexes.]  The words are reassembled to produce a regex
        matching the log line, and a counter is maintained for each regex.
        Contiguous sequences of \texttt{.+} in the newly assembled regexes
        are collapsed to a single \texttt{.+}; any resulting duplicate
        regexes and their counters are combined.  If a regex's counter is
        lower than the threshold supplied by the user the regex is
        discarded; this second threshold is independent of the threshold
        used to differentiate between fixed and variable words, but once
        again 1\% of log lines is a good starting point.

    \item [Test the new regexes.]  The final step replaces keywords in the
        new regexes in the same way as \parsername{} does, and compiles
        each regex to check they are valid.  A match will be attempted
        between all of the new regexes and each of the original unparsed
        log lines (not the pre-processed log lines); the user will be
        warned if a log line is not matched by any regex, or if a log line
        is matched by more than one regex.  The number of log lines each
        regex matches is counted, as is the number of log lines matched by
        all regexes, though a log line is counted once only, even if
        matched by more than one regex.

@@

The new regexes are displayed for the user to add to the ruleset, either as
new rules or merged into the regexes of existing rules; also displayed are
the number of unrecognised log lines each regex was expected to match, and
the number it actually matched, to help the user notice problems in the new
regexes.  Discarding regexes will result in some of the unrecognised log
lines not being matched; when the ruleset has been augmented with the new
regexes, the original log files should be parsed again, and any remaining
unparsed log lines used as input to this utility.

This utility is not expected to create perfect regexes, but it greatly
reduces the effort required to deal with unrecognised log lines.  The
regexes it generates will be self-contained: a parser that relies on using
cascaded parsing would require a modified algorithm, perhaps by replacing
the pre-processing stage with one that applies existing cascading rules to
each log line, and uses the resulting modified log lines for the remainder
of the algorithm.

@@

Each rule's regex will have keywords expanded when the ruleset is loaded,
for several reasons:

@@

    \item It simplifies both reading and writing of regexes and helps to
        make each regex largely self-documenting.  For example, the meaning
        of \_\_CLIENT\_HOSTNAME\_\_ is immediately clear, whereas its
        expansion \verb!(?:unknown|(?:[-.\w]+))! needs to be deciphered
        each time it is encountered.

    \item It avoids needless repetition of complex regex components, and
        allows the components to be corrected or improved in one location.
        For example, \_\_SENDER\_\_ is used in 68 rules; if a mistake is
        discovered in it the mistake only needs to be corrected in one
        place.

    \item It enables automatic extraction and saving of captured data.  The
        regex snippets use Perl 5.10's named capture buffers~\cite{perlre}
        to capture data, so the mapping between captures and fields does
        not need to be explicitly specified by the rule.

@@

To improve efficiency, the keywords are expanded and every rule's regex is
compiled before attempting to parse the log file, otherwise every regex
would be recompiled each time it was used, resulting in a large, data
dependent slowdown, as described in \sectionref{Caching compiled regexes}.
Most of the keywords are named after the fields in the connections or
results tables they populate: \_\_CLIENT\_HOSTNAME\_\_, \_\_CLIENT\_IP\_\_,
\_\_DELAY\_\_, \_\_DELAYS\_\_, \_\_ENHANCED\_STATUS\_CODE\_\_,
\_\_HELO\_\_, \_\_MESSAGE\_ID\_\_, \newline{} \_\_QUEUEID\_\_,
\_\_RECIPIENT\_\_, \_\_SENDER\_\_, \_\_SERVER\_HOSTNAME\_\_,
\_\_SERVER\_IP\_\_, \_\_SIZE\_\_, and \_\_SMTP\_CODE\_\_.

@@

Some similarity exists between regex components and cascaded parsing: each
regex component resembles a rule that recognises part of a log line and
consumes it, leaving the remainder to be recognised by other components or
cascaded rules.  The major difference between the two is that regex
components are explicitly used by the author of the ruleset, whereas
cascaded parsing would be dynamically applied by the framework.

@@

    \item Sort rules by program and regex e.g.\ with an \acronym{SQL} query
        similar to:                                     \newline{}
        \verb!SELECT program, regex!                    \newline{}
        \verb!    FROM rules!                           \newline{}
        \verb!    ORDER BY program, regex;!             \newline{}
        The rules are sorted first by program, then by regex, because rules
        cannot overlap if their programs are different.  Note that this
        query does not properly deal with generic rules whose program is
        \texttt{*}; those rules will be used on all log lines that have not
        been recognised by program-specific rules.

        Inspect the list to see if any pair of regexes look suspiciously
        similar; overlapping regexes will often lie beside one another when
        sorted.

@@

\section{Limitations And Possible Improvements}

@@

This chapter has presented \parsername{}, the parser implemented for this
project, beginning with the assumptions made during its development.  A
simplified flowchart shows the most common paths taken through Postfix and
\parsername{}, accompanied by a description of the stages and transitions.
A database provides storage for rules and for data gathered from log files;
any further use of that data is dependent on a clear understanding of the
database schema, so the role of the database schema as an \acronym{API} is
described, followed by a diagram and a detailed description of the database
schema.  The framework is documented next, including the steps it takes
during initialisation, the parsing process, and the conveniences it offers
to users.  The performance data collected by the framework is described, as
are the ways in which various optimisations can be disabled to demonstrate
their effect, and the debugging options the framework provides.  The
implementation of actions in \parsername{} is documented, including how
frequently each action is specified by rules, and brief descriptions of why
some actions are more popular than others.  All the actions that are part
of \parsername{} are described in detail, followed by the process of adding
a new action to the parser.  The final component of the architecture, the
rules, is also the most visible component, and its implementation is
examined in detail with: a sample rule and an explanation of how every
field in that rule is used; a description of how to add new rules and
determine the values that should be used for each field; an explanation of
the algorithm used by the utility that creates new regexes from unparsed
log lines, and how it differs from the original algorithm it is based on.
How \parsername{} uses rule conditions and overlapping rules is discussed,
accompanied by a description of the regex snippets the framework provides
to simplify the regexes used in rules.  The many complications and
difficulties encountered while writing \parsername{}, and the solutions
developed to overcome them, are documented in depth; also documented are
how solutions interact, and which action or actions each solution is
implemented in.  This chapter concludes with a list of the limitations
identified in \parsername{}.  The next chapter will evaluate this
implementation, examining both \parsernames{} efficiency and the coverage
it achieves when parsing log files.

>>>>

The next three sections document the implementation of the three components
of the architecture.  First is the framework, including its initialisation
phase, the parsing process, and the conveniences it offers to users of
\parsername{}; that section finishes by describing features that are
important for this thesis but not strictly necessary for parsing: the
performance data collected by the framework, the optimisations that can be
disabled to show their effect, and the debugging options the framework
provides.  The second component of the architecture is the actions,
starting with a graph showing how often each action is specified by rules,
a description of why some actions are more popular than others, and how
this popularity has no influence on how often an action is invoked.  Every
action that is part of \parsername{} is documented, and the actions section
concludes with a description of the process of adding a new action.  Rules
are the third component of the architecture, and also the most visible to
advanced users, e.g.\ systems administrators, because it is likely that
they will need to add their own rules to recognise their own log lines.  A
sample rule used by \parsername{} is examined, with every field clearly
documented; that is followed by a description of adding new rules, and how
to determine the value of each of the new rule's fields.  \parsername{}
provides a utility to create \regexes{} from unrecognised log lines; the
algorithm it uses is documented, including the differences between it and
the original algorithm it is based on.  The rules section finishes with a
discussion of how \parsername{} uses rule conditions and overlapping rules,
plus a description of the \acronym{regex} snippets provided to ease the
process of writing \regexes{}.

@@

The database schema can be conceptually divided in two: the rules used to
recognise log lines, and the data saved from the parsing of log files.
Each rule has a \acronym{regex} to recognise log lines and capture data
from them, and specifies the action to be invoked when a log line is
recognised; they also have several other fields used by the parser, and
several fields that aid the user in understanding the meaning of the log
lines recognised by each rule.  The rules are described in detail in
\sectionref{rules in implementation}, but the rules table is documented in
\sectionref{rules table} with the rest of the database schema.

@@

\subsection{Using a Database to Provide an Application Programming
Interface}

@@

    \item [regex] The \acronym{regex} to recognise log lines with, as
        documented in \sectionref{regex components}.

@@

    \item That the \acronym{regex} is valid.  The \acronym{regex} will have
        \acronym{regex} components expanded (see \sectionref{regex
        components}), and will also be compiled for efficiency
        (\sectionref{Caching compiled regexes}).

    \item For overlap between the data captured by the \acronym{regex} and
        additional data specified in either \texttt{connection\_data} or
        \texttt{result\_data}.

    \item That \texttt{connection\_data}, \texttt{result\_data}, and the
        \acronym{regex} captures specify valid fields to save data to.

@@

    \item Each \acronym{regex} is compiled and the result cached when the
        ruleset is loaded; this optimisation can be disabled and the
        \acronym{regex} compiled every time it is used when recognising log
        lines.  The effect this optimisation has is addressed shown in
        \sectionref{Caching compiled regexes}.

@@

    \item Every time a log line is recognised, the recognising rule's
        \acronym{regex} and the log line can be printed, so that the user
        can verify that the correct rule recognised each log line.  This is
        equivalent to setting every rule's debug field to true.

@@

\subsection{Description of Each Action}

@@

If the log line has enough information to identify the correct connection
or mail, each action will save all the data captured by the recognising
rule's \acronym{regex} to the connection or mail; usually, log lines that
lack identifying information will be processed by the
\action{UNINTERESTING} action.  Each action is passed the same arguments:

@@

    \item [data] The data captured from the log line by the rule's
        \acronym{regex}.

@@

    \item [CLEANUP\_PROCESSING] \daemon{cleanup} processes every mail that
        passes through Postfix; details of what it does are available in
        \sectionref{Postfix Daemons}.  This action saves all data captured
        by the rule's \acronym{regex} if the log line has not come after a
        timeout (see \sectionref{discarding cleanup log lines}); it also
        creates the mail if necessary, setting its unknown origin flag (see
        \sectionref{pickup logging after cleanup}).

@@

    \item [DELIVERY\_REJECTED] Postfix rejected a mail delivery attempt
        from the remote client.  This is the action most frequently
        specified by rules, because so many different restrictions are used
        to reject delivery attempts.  This action is quite simple: if the
        log line contains a queueid, save the data captured by the rule's
        \acronym{regex} to the mail identified by that queueid; otherwise
        save it to the connection identified by the \acronym{pid} in the
        log line.

@@

    \item [MAIL\_BOUNCED] This action behaves in exactly the same way as
        the \action{SAVE\_DATA} action; it saves all data captured by the
        recognising rule's \acronym{regex}, and does nothing more.  It is a
        separate action to distinguish delivery attempts that bounce from
        other delivery attempts.

@@

    \item [MAIL\_SENT] This action behaves in exactly the same way as the
        \action{SAVE\_DATA} action; it saves all data captured by the
        recognising rule's \acronym{regex}, and does nothing more.  It is a
        separate action to distinguish successful delivery attempts from
        other delivery attempts.

@@

    \item [SAVE\_DATA] Every action that can locate the correct data
        structure in the state tables saves any data captured by the
        recognising rule's \acronym{regex} to it.  The \action{SAVE\_DATA}
        saves data in this way but does not do anything else; it is invoked
        for log lines that contain useful data but do not require any
        further processing.

@@

The role of the rules in the architecture is covered in detail in
\sectionref{rules in architecture}; this section is concerned with the
practical aspects of how rules are implemented and used in \parsername{}.
The structure of the rules table has already been documented in
\sectionref{rules table}; that description will not be repeated here, but
should be fleshed out by the sample rule in \sectionref{example rule in
implementation}.  The process of creating new rules from unparsed log lines
is dealt with in \sectionref{creating new rules in implementation},
followed by the algorithm used by the utility supplied with \parsername{}
that creates \regexes{} from unparsed log lines; the \acronym{regex}
components provided by \parsername{} to ease writing of complex \regexes{}
are covered in \sectionref{regex components}.  The rule conditions used in
\parsername{} are the penultimate topic (\sectionref{rule conditions in
implementation}), and this section concludes with some suggestions for how
to detect overlapping rules.

@@

    \item [program and regex:] program is used to restrict the log lines
        this rule will be used to recognise; see \sectionref{rule
        conditions in implementation} for details.  The \acronym{regex}
        does the actual recognition of log lines, and data captured by the
        \acronym{regex} (e.g.\ sender, recipient) will be automatically
        saved to the results and connections tables.  The \acronym{regex}
        components used in the \acronym{regex} are described in
        \sectionref{regex components}.

@@

To add a new rule a new row must be added to the rules table in the
database, containing all the required attributes: action, name,
description, program, and \acronym{regex}; all the other attributes are
either optional, are set by the parser, or have sensible defaults.  The
name and description fields should be set based on the meaning of the log
line, to help others understand which log lines this rule will recognise;
the value for the program field will be obvious from the unrecognised log
lines.  The action depends on the what the log line represents, e.g.\ a
delivery rejection, a mail being delivered, some useful information, or
something else; examine the list of actions in \sectionref{actions in
detail in implementation} to determine the correct one.  The
\acronym{regex} needs to be based on the log line, but see below for a tool
to ease the process of creating \regexes{} from unrecognised log lines.

@@

A program is provided with \parsername{} to ease the process of creating
new rules from unrecognised log lines, based on the algorithm developed by
Risto Vaarandi for his \acronym{SLCT}~\cite{slct-paper}.  The differences
between the two algorithms will be outlined as part of the explanation
below.  The core of the \acronym{SLCT} algorithm is quite simple: programs
generally create log lines by substituting variable words into a fixed
pattern, and analysis of the frequency with which each word occurs can be
used to determine whether the word is variable or part of the fixed
pattern.  This classification can be used to group similar log lines and
generate a \acronym{regex} to match each group of log lines.  The algorithm
has five steps:

@@

    \item [Pre-process the input.]  The modified algorithm begins by
        leveraging the knowledge gained from writing rules and developing
        \parsername{}; it performs many substitutions on the input log
        lines, replacing commonly occurring variable terms (e.g.\ email
        addresses, IP addresses, the standard start of rejection messages)
        with keywords described in \sectionref{regex components}.  The
        purpose of this step is to utilise existing knowledge to create
        more accurate \regexes{}; it replaces many variable words with
        fixed words, improving the subsequent classification of words as
        fixed or variable.  Regex metacharacters in the log line will be
        escaped, to avoid generating invalid or incorrect \regexes{}.  The
        altered log lines are written to a temporary file, which the next
        stage will use instead of the original input file.

@@

    \item [Classify words based on their frequency.]  The frequency of each
        \texttt{(word, word's position within the log line)} tuple is
        checked: if its frequency is greater than the threshold supplied by
        the user (1\% of all log lines is generally a good starting point),
        it is classified as a fixed word, otherwise it is classified as a
        variable term.  If a variable term appears sufficiently often it
        will be misclassified as a fixed term, but that should be noticed
        by the user when reviewing the new \regexes{}, and will be obvious
        when the new rules do not recognise some log lines they are
        expected to.  Variable terms are replaced by \texttt{.+} to match
        one or more of any character.

    \item [Build regexes.]  The words are reassembled to produce a
        \acronym{regex} matching the log line, and a counter is maintained
        for each \acronym{regex}.  Contiguous sequences of \texttt{.+} in
        the newly assembled \regexes{} are collapsed to a single
        \texttt{.+}; any resulting duplicate \regexes{} and their counters
        are combined.  If a \acronym{regex}'s counter is lower than the
        threshold supplied by the user the \acronym{regex} is discarded;
        this second threshold is independent of the threshold used to
        differentiate between fixed and variable words, but once again 1\%
        of log lines is a good starting point.

    \item [Test the new regexes.]  The final step replaces keywords in the
        new \regexes{} in the same way as \parsername{} does, and compiles
        each \acronym{regex} to check they are valid.  A match will be
        attempted between all of the new \regexes{} and each of the
        original unparsed log lines (not the pre-processed log lines); the
        user will be warned if a log line is not matched by any
        \acronym{regex}, or if a log line is matched by more than one
        \acronym{regex}.  The number of log lines each \acronym{regex}
        matches is counted, as is the number of log lines matched by all
        \regexes{}, though a log line is counted once only, even if matched
        by more than one \acronym{regex}.

@@

The new \regexes{} are displayed for the user to add to the ruleset, either
as new rules or merged into the \regexes{} of existing rules; also
displayed are the number of unrecognised log lines each \acronym{regex} was
expected to match, and the number it actually matched, to help the user
notice problems in the new \regexes{}.  Discarding \regexes{} will result
in some of the unrecognised log lines not being matched; when the ruleset
has been augmented with the new \regexes{}, the original log files should
be parsed again, and any remaining unparsed log lines used as input to this
utility.

This utility is not expected to create perfect \regexes{}, but it greatly
reduces the effort required to deal with unrecognised log lines.  The
\regexes{} it generates will be self-contained: a parser that relies on
using cascaded parsing would require a modified algorithm, perhaps by
replacing the pre-processing stage with one that applies existing cascading
rules to each log line, and uses the resulting modified log lines for the
remainder of the algorithm.

@@

Each rule's \acronym{regex} will have keywords expanded when the ruleset is
loaded, for several reasons:

@@

    \item It simplifies both reading and writing of \regexes{} and helps to
        make each \acronym{regex} largely self-documenting.  For example,
        the meaning of \_\_CLIENT\_HOSTNAME\_\_ is immediately clear,
        whereas its expansion \verb!(?:unknown|(?:[-.\w]+))! needs to be
        deciphered each time it is encountered.

    \item It avoids needless repetition of complex \acronym{regex}
        components, and allows the components to be corrected or improved
        in one location.  For example, \_\_SENDER\_\_ is used in 68 rules;
        if a mistake is discovered in it the mistake only needs to be
        corrected in one place.

    \item It enables automatic extraction and saving of captured data.  The
        \acronym{regex} snippets use Perl 5.10's named capture
        buffers~\cite{perlre} to capture data, so the mapping between
        captures and fields does not need to be explicitly specified by the
        rule.

@@

To improve efficiency, the keywords are expanded and every rule's
\acronym{regex} is compiled before attempting to parse the log file,
otherwise every \acronym{regex} would be recompiled each time it was used,
resulting in a large, data dependent slowdown, as described in
\sectionref{Caching compiled regexes}.  Most of the keywords are named
after the fields in the connections or results tables they populate:
\_\_CLIENT\_HOSTNAME\_\_, \_\_CLIENT\_IP\_\_, \_\_DELAY\_\_,
\_\_DELAYS\_\_, \_\_ENHANCED\_STATUS\_CODE\_\_, \_\_HELO\_\_,
\_\_MESSAGE\_ID\_\_, \newline{} \_\_QUEUEID\_\_, \_\_RECIPIENT\_\_,
\_\_SENDER\_\_, \_\_SERVER\_HOSTNAME\_\_, \_\_SERVER\_IP\_\_, \_\_SIZE\_\_,
and \_\_SMTP\_CODE\_\_.

@@

Some similarity exists between \acronym{regex} components and cascaded
parsing: each \acronym{regex} component resembles a rule that recognises
part of a log line and consumes it, leaving the remainder to be recognised
by other components or cascaded rules.  The major difference between the
two is that \acronym{regex} components are explicitly used by the author of
the ruleset, whereas cascaded parsing would be dynamically applied by the
framework.

@@

    \item Sort rules by program and \acronym{regex} e.g.\ with an
        \acronym{SQL} query similar to:                 \newline{}
        \verb!SELECT program, regex!                    \newline{}
        \verb!    FROM rules!                           \newline{}
        \verb!    ORDER BY program, regex;!             \newline{}
        The rules are sorted first by program, then by \acronym{regex},
        because rules cannot overlap if their programs are different.  Note
        that this query does not properly deal with generic rules whose
        program is \texttt{*}; those rules will be used on all log lines
        that have not been recognised by program-specific rules.

        Inspect the list to see if any pair of \regexes{} look suspiciously
        similar; overlapping \regexes{} will often lie beside one another
        when sorted.

@@

% Move section start onto next page because extra spacing causes it to be
% badly positioned.
\newpage{}

\section{Limitations and Possible Improvements}

@@

This chapter has presented \parsername{}, the parser implemented for this
project, beginning with the assumptions made during its development.  A
simplified flowchart shows the most common paths taken through Postfix and
\parsername{}, accompanied by a description of the stages and transitions.
A database provides storage for rules and for data gathered from log files;
any further use of that data is dependent on a clear understanding of the
database schema, so the role of the database schema as an \acronym{API} is
described, followed by a diagram and a detailed description of the database
schema.  The framework is documented next, including the steps it takes
during initialisation, the parsing process, and the conveniences it offers
to users.  The performance data collected by the framework is described, as
are the ways in which various optimisations can be disabled to demonstrate
their effect, and the debugging options the framework provides.  The
implementation of actions in \parsername{} is documented, including how
frequently each action is specified by rules, and brief descriptions of why
some actions are more popular than others.  All the actions that are part
of \parsername{} are described in detail, followed by the process of adding
a new action to the parser.  The final component of the architecture, the
rules, is also the most visible component, and its implementation is
examined in detail with: a sample rule and an explanation of how every
field in that rule is used; a description of how to add new rules and
determine the values that should be used for each field; an explanation of
the algorithm used by the utility that creates new \regexes{} from unparsed
log lines, and how it differs from the original algorithm it is based on.
How \parsername{} uses rule conditions and overlapping rules is discussed,
accompanied by a description of the \acronym{regex} snippets the framework
provides to simplify the \regexes{} used in rules.  The many complications
and difficulties encountered while writing \parsername{}, and the solutions
developed to overcome them, are documented in depth; also documented are
how solutions interact, and which action or actions each solution is
implemented in.  This chapter concludes with a list of the limitations
identified in \parsername{}.  The next chapter will evaluate this
implementation, examining both \parsernames{} efficiency and the coverage
it achieves when parsing log files.

### 2009-08-27 doc/logparser-other-parsers-review.tex
Turned regex into an acronym everywhere.

\acronym{LMA} requires major changes to the parser code to parse new log
lines or to extract additional data.  The code is structured as a long
series of blocks that each handle all log lines matching a single
\acronym{regex}, so parsing new log lines requires modifying an existing
regex or carefully inserting a new block in the correct place; extracting
extra data would require modifying multiple blocks, regexes, or both.
\regexglossarydescription{}

@@

It also counts the number of log lines parsed and skipped.  It omits client
IP address and hostname, server IP address, HELO hostname, queueid,
timestamps of each log line, enhanced status code, and message-id\@.  Log
lines are correlated based on the queueid (called sessionname [sic] within
Logrep), but not by \acronym{pid}.  The parsing is error prone: empty
fields are saved when the log line does not match the regex, though it
appears that they will not overwrite existing data.  Most notably, rejected
delivery attempts are completely ignored, making it unsuitable for the
purposes of this project.  It does not try to address any of the
complications in \sectionref{complications} except for correlating log
lines by queueid.

>>>>

\acronym{LMA} requires major changes to the parser code to parse new log
lines or to extract additional data.  The code is structured as a long
series of blocks that each handle all log lines matching a single
\acronym{regex}, so parsing new log lines requires modifying an existing
\acronym{regex} or carefully inserting a new block in the correct place;
extracting extra data would require modifying multiple blocks, \regexes{},
or both.  \regexglossarydescription{}

@@

It also counts the number of log lines parsed and skipped.  It omits client
IP address and hostname, server IP address, HELO hostname, queueid,
timestamps of each log line, enhanced status code, and message-id\@.  Log
lines are correlated based on the queueid (called sessionname [sic] within
Logrep), but not by \acronym{pid}.  The parsing is error prone: empty
fields are saved when the log line does not match the \acronym{regex},
though it appears that they will not overwrite existing data.  Most
notably, rejected delivery attempts are completely ignored, making it
unsuitable for the purposes of this project.  It does not try to address
any of the complications in \sectionref{complications} except for
correlating log lines by queueid.
